---
title: "In-class_Ex04 : Geographically Weighted Summary Statistics - gwModel methods"
author: "Kalpit Kulshrestha"
date: "September 20, 2025"
date-modified: "last-modified"
execute: 
  eval: true
  echo: true
  warning: false
  freeze: true
---

## Loading the package

In this in-class exercise, sf, spdep, tmap, tidyverse, knitr and GWmodel will be used.

### The code

```{r}
pacman::p_load(sf, ggstatsplot, tmap, tidyverse, knitr, GWmodel)
```

## Preparing the Data

For this in-class exercise, Hunan shapefile and Hunan_2012 data file will be used.

Using the steps you learned from previous hands-on, complete the following tasks:

-   import Hunan shapefile and parse it into a sf polygon feature object.
-   import Hunan_2012.csv file parse it into a tibble data.frame.
-   join Hunan and Hunan_2012 data.frames.

### Importing Hunan shapefile

```{r}
hunan_sf <- st_read(dsn = "data/geospatial", 
                 layer = "Hunan")
```

### Importing Hunan_2012 table

```{r}
hunan2012 <- read_csv("data/aspatial/Hunan_2012.csv")
```

### Joining Hunan and Hunan_2012

```{r}
hunan_sf <- left_join(hunan_sf, hunan2012) %>%
  select(1:3, 7, 15, 16, 31, 32)
```

## Mapping GDPPC

::::: panel-tabset
Using the steps you learned from Hands-on Exercise 5, prepare a choropleth map showing the geographic distribution of GDPPC of Hunan Province.

```{r}
#| fig-width: 12
#| fig-height: 6
#| echo: false
basemap <- tm_shape(hunan_sf) +
  tm_polygons() +
  tm_text("NAME_3", size=0.5)

gdppc <- qtm(hunan_sf, "GDPPC")
tmap_arrange(basemap, gdppc, asp=1, ncol=2)
```

### The code

```{r}
#| fig-width: 12
#| fig-height: 8
#| eval: false
basemap <- tm_shape(hunan_sf) +
  tm_polygons() +
  tm_text("NAME_3", size=0.5)

gdppc <- qtm(hunan_sf, "GDPPC")
tmap_arrange(basemap, gdppc, asp=1, ncol=2)
```

## Converting to SpatialPolygonDataFrame

::: callout-note
GWmodel presently is built around the older sp and not sf formats for handling spatial data in R.
:::

```{r}
hunan_sp <- hunan_sf %>%
  as_Spatial()
```

## Geographically Weighted Summary Statistics with adaptive bandwidth

### Determine adaptive bandwidth

::: panel-tabset
### Cross-validation

```{r}
bw_CV <- bw.gwr(GDPPC ~ 1, 
             data = hunan_sp,
             approach = "CV",
             adaptive = TRUE, 
             kernel = "bisquare", 
             longlat = T)
bw_CV
```
:::

### AIC

```{r}
bw_AIC <- bw.gwr(GDPPC ~ 1, 
             data = hunan_sp,
             approach ="AIC",
             adaptive = TRUE, 
             kernel = "bisquare", 
             longlat = T)
bw_AIC
```

## Geographically Weighted Summary Statistics with adaptive bandwidth

### Computing geographically wieghted summary statistics

```{r}
gwstat <- gwss(data = hunan_sp,
               vars = "GDPPC",
               bw = bw_AIC,
               kernel = "bisquare",
               adaptive = TRUE,
               longlat = T)
```

### Preparing the output data

Code chunk below is used to extract **SDF** data table from **gwss** object output from `gwss()`. It will be converted into data.frame by using `as.data.frame()`.

```{r}
gwstat_df <- as.data.frame(gwstat$SDF)
```

```{r}
hunan_gstat <- cbind(hunan_sf, gwstat_df)
```
:::::

```{r}
#| fig-height: 9
#| echo: false
tm_shape(hunan_gstat) +
  tm_fill("GDPPC_LM",
          n = 5,
          style = "quantile") +
  tm_borders(alpha = 0.5) +
  tm_layout(main.title = "Distribution of geographically wieghted mean",
            main.title.position = "center",
            main.title.size = 2.0,
            legend.text.size = 1.2,
            legend.height = 1.50, 
            legend.width = 1.50,
            frame = TRUE)
```

### The code

```{r}
#| eval: false
#| include: false
tm_shape(hunan_gstat) +
  tm_fill("GDPPC_LM",
          n = 5,
          style = "quantile") +
  tm_borders(alpha = 0.5) +
  tm_layout(main.title = "Distribution of geographically wieghted mean",
            main.title.position = "center",
            main.title.size = 2.0,
            legend.text.size = 1.2,
            legend.height = 1.50, 
            legend.width = 1.50,
            frame = TRUE)
```
