---
title: "In-class Exercise 10"
author: "Kalpit Kulshrestha"
---
# Overview
Market area analysis (also known as Trade Area Analysis) is the process of defining and studying the geographical area from which a business draws its customers. It involves analyzing data on consumer demographics, spending habits, and competitors to understand market viability, identify opportunities, and inform decisions about location, marketing, and expansion. It often combines empirical observation, like customer spotting, with mathematical models to identify market segments and understand customer behavior.

Huff model is a geospatial analysis method commonly used in market area analysis to predict the probability that a consumer from a specific area will patronize a particular store, based on the store’s attractiveness and distance relative to competitors. It calculates a store’s market share by considering factors like size or sales for attractiveness and distance or travel time, with the probability decreasing as distance or competition increases. This tool helps businesses assess potential revenue, define trade areas, and make informed decisions about site selection

## Setting the scene
A fast-food franchise is a business model in which an individual (the franchisee) acquires the rights to operate a restaurant under an established brand owned by a franchisor. This arrangement typically involves franchise fees and adherence to standardized operational procedures. Key considerations prior to investment include the scale of initial capital outlay, the strength of the brand, and the strategic importance of store location. Capital investment requirements vary significantly, generally ranging from approximately $200,000 to more than $2 million. Major global fast-food franchise brands include McDonald’s, Burger King, Pizza Hut, KFC, and Taco Bell.

In this exercise, you will assume the role of a data analyst within the business development team of a fast-food franchise. Your objective is to delineate the trade areas of existing stores in Taichung City and to produce corresponding spatial maps to support market analysis and strategic decision-making.

## Getting started - Installing packages
```{r}
pacman::p_load(sf, tidyverse, MCI, tmap, ggstatsplot)
```

## importing data

### Geospatial data
```{r}
stores <- st_read(
  dsn = "data/geospatial",
  layer = "STORES_TC")
```
```{r}
town <- st_read(
  dsn = "data/geospatial",
  layer = "TOWN_TC")
```
```{r}
village <- st_read(
  dsn = "data/geospatial",
  layer = "VILLAGE_TC")
```
```{r}
tmap_mode("view")
tm_shape(town) +
  tm_polygons() +
tm_shape(village) +
  tm_polygons() + 
tm_shape(stores) +
  tm_dots()
```
```{r}
tmap_mode("plot")
```
### Aspaital data
```{r}
pop2019 <- read_csv(
  "data/aspatial/pop2019.csv")

store_sales <- read_csv(
  "data/aspatial/Store_sales_data.csv")
```
```{r}
summary(store_sales)
```
## Data Prepration

### Aggregating store sales
```{r}
store_sales_aggre <- store_sales %>%
  group_by(`Store Code`) %>%
  summarise(
    "Total_Bill" = sum(Bills, na.rm = TRUE),
    "Total_Sales" = sum(Amount, na.rm = TRUE))
```


### Relational join
```{r}
stores <- stores %>%
  left_join(store_sales_aggre,
            by = join_by(
              Store_CD == `Store Code`)) %>%
  mutate(Store_size = 100)

```

```{r}
summary(stores)
```
```{r}

```
### Tidying pop2019 data.frame
```{r}
glimpse(pop2019)
```
```{r}
glimpse(village)
```
```{r}
pop2019 <- pop2019 %>%
  mutate(V_ID = str_remove_all(
    V_ID, "-"))
```

```{r}
village <- village %>%
  left_join(pop2019,
            by = join_by(
              VILLCODE == V_ID))
```

### Check for missing or zero values
```{r}
summary(village)
```
```{r}
village <- village %>%
  filter(!is.na(P_CNT)) 
```

```{r}
any(is.na(village$H_CNT))
```
```{r}
any(is.na(village$P_CNT))
```
```{r}
any(is.na(village$M_CNT))
```
```{r}
any(is.na(village$F_CNT))
```
## Computing distance Matrix
```{r}
village_center <- st_point_on_surface(village)
distmat <- st_distance(village_center, stores) 
```

```{r}
rownames(distmat) <- village$`VILLCODE`
colnames(distmat) <- stores$`Store_CD`

distmat_long <- distmat %>%
  as.data.frame() %>%
  rownames_to_column(var = "VILLCODE") %>%
  pivot_longer(
    cols = -VILLCODE,
    names_to = "Store_CD",
    values_to = "distance"
  ) %>%
  mutate(distance = as.numeric(distance)) 
```

## Preparing Internaction matrix
```{r}
interaction_matrix <- distmat_long %>%
  left_join(
    stores %>% 
      st_drop_geometry() %>% 
      select(Store_CD, Store_size),
    by = "Store_CD") %>%
  relocate(
    distance, .after = last_col())
```
```{r}
summary(interaction_matrix)
```
## Computing Huff model
```{r}
huff_share <- huff.shares(
  interaction_matrix, "VILLCODE", "Store_CD", 
  "Store_size", "distance")
```

### Trade area analysis for a single store
First, let us select market share data of a store by usign Store_CD field (i.e. TG).
```{r}
market_share <- huff_share %>%
  filter(Store_CD == "TG")
```
Next, trade areas of the selected will be derived by using the code chun below.
```{r}
Trade_area <- village %>%
  left_join(market_share)
```
Before plotting the trade area map, code chunk below is used to extract the selected store from stores sf data.frame.
```{r}
selected_store <- stores %>%
  filter(Store_CD == "TG")
```

```{r}
tm_shape(Trade_area) +
  tm_polygons(fill = "p_ij",
              fill.scale = tm_scale_intervals(
                style = "quantile", 
                n = 10,
                values = "brewer.blues"),
              fill.legend = tm_legend(
                title = "Patronage Probability")) +
  tm_shape(selected_store) +
    tm_dots(size = 0.3,
            fill = "red",
            col = "black") +
  tm_title("Distribution of Patronage Probability of Store TG by village") +
  tm_layout(frame = TRUE) +
  tm_borders(fill_alpha = 0.5) +
  tm_compass(type="8star", size = 2) +
  tm_scalebar() +
  tm_grid(alpha =0.2) 
```

### store sales analysis
```{r}
huff_share_pp <- huff_share %>%
  left_join(village %>% 
      st_drop_geometry() %>% 
      select(VILLCODE, P_CNT),
    by = "VILLCODE") %>%
  relocate(
    distance, .after = last_col())
```

```{r}
huff_total <- shares.total(
  huff_share_pp, "VILLCODE", "Store_CD",
  "p_ij", "P_CNT")
```

```{r}
store_sales_est <- stores %>%
  left_join(huff_total,
            by = join_by(
              "Store_CD" == "suppliers_single"))
```

```{r}
store_sales_nongeom <- store_sales_est %>% st_drop_geometry()

ggscatterstats(
  store_sales_nongeom,
  x = sum_E_j,
  y = Total_Sales)
```
```{r}
huff_share_wider <- huff_share %>%
  select(VILLCODE, Store_CD, p_ij) %>%
    pivot_wider(names_from = Store_CD, 
                values_from = p_ij)
```

