---
title: "Hands-on Exercise 01A"
author: "Kalpit Kulshrestha"
---

# Geospatial Data Wrangling with R

Geospatial Data Science is a process of importing, wrangling, integrating, and processing geographically referenced data sets. In this hands-on exercise, you will learn how to perform geospatial data science tasks in R by using sf package.

```{r}
pacman::p_load(sf, tidyverse)
```

## Importing the data

### Importing Polygon feature data for Masterplan

```{r}
mpsz = st_read(dsn = "K:/kalpitkulshrestha24/ISSS626/Hands-on_Exercise/Hands-on_EX01/data/geospatial/MasterPlan2014SubzoneBoundaryWebSHP", 
                  layer = "MP14_SUBZONE_WEB_PL")

```

```{r}
st_crs(mpsz)
```

### Importing polyline feature data in shapefile form for Cycling

```{r}
Cyclingpath = st_read(dsn = "K:/kalpitkulshrestha24/ISSS626/Hands-on_Exercise/Hands-on_EX01/data/geospatial/CyclingPath_Apr2025", layer = "CyclingPathGazette")

```

```{r}
st_crs(Cyclingpath)
```

```{r}
Cyclingpath <- st_transform(Cyclingpath, crs = 3414)
mpsz  <- st_transform(mpsz,  crs = 3414)
```

### Importing polyline feature data in shapefile form for Preschoolslocation dataset

```{r}
preschool = st_read("K:/kalpitkulshrestha24/ISSS626/Hands-on_Exercise/Hands-on_EX01/data/geospatial/PreSchoolslocation/PreSchoolslocation.kml")
```

## A simple feature dataframe

### Retriving geometery list-column

```{r}
st_geometry(mpsz)
```

```{r}
st_geometry(preschool)
```

```{r}
st_geometry(Cyclingpath)
```

### Using glimpse() funtion

```{r}
glimpse(mpsz)
```

### Using head() funtion

```{r}
head(mpsz)
```

## ploting the Geospatial data

```{r}
plot(mpsz)
```

```{r}
plot(st_geometry(mpsz))
plot(mpsz["PLN_AREA_N"])
```

```{r}
plot(st_geometry(mpsz))
plot(st_geometry(preschool), 
     add = TRUE)
```

## Working with Projection

### projection transformation

```{r}
st_crs(mpsz)
```

```{r}
mpsz <- st_set_crs(mpsz, 3414)
```

```{r}
st_crs(mpsz)
```

### Transforming Projection

```{r}
preschool <- st_transform(preschool, 
                              crs = 3414)
st_geometry(preschool)
plot(st_geometry(mpsz))
plot(st_geometry(preschool), add = TRUE)
```

## Importing and Converting An Aspatial Data

```{r}
listings <- read_csv("K:/kalpitkulshrestha24/ISSS626/Hands-on_Exercise/Hands-on_EX01/data/aspatial/listings.csv")
list(listings)
```

```{r}
listings_sf <- st_as_sf(listings, 
                       coords = c("longitude", "latitude"),
                       crs=4326) %>%
  st_transform(crs = 3414)
```

```{r}
glimpse(listings_sf)
plot(st_geometry(mpsz))
plot(st_geometry(listings_sf), add = TRUE)
```

## Geoprocessing with sf package

### Use Case 1 : Land Aquition analysis

```{r}
buffer_cycling <- st_buffer(
  Cyclingpath, dist=5, nQuadSegs = 30)
```

```{r}
buffer_cycling$AREA <- st_area(buffer_cycling)
```

```{r}
buffer_cycling <- buffer_cycling %>%
  mutate(AREA = st_area(geometry))
```

```{r}
sum(buffer_cycling$AREA)
```

```{r}
mpsz_selected <- mpsz %>% filter(SUBZONE_N == "TAMPINES WEST") 
```

```{r}
st_crs(buffer_cycling)
st_crs(mpsz_selected)
```

```{r}
buffer_cycling_selected <- st_intersection(
  buffer_cycling, mpsz_selected)
```

```{r}
plot(buffer_cycling_selected)
```

### Use Case 2: To determine the number of pre-schools by planning subzone

```{r}
mpsz$`PreSch Count`<- lengths(st_intersects(mpsz, preschool))
```

```{r}
summary(mpsz$`PreSch Count`)
```

```{r}
top_n(mpsz, 1, `PreSch Count`)
```

```{r}
mpsz$Area <- mpsz %>%
  st_area()
```

```{r}
mpsz <- mpsz %>%
  mutate(`PreSch Density` = `PreSch Count`/Area * 1000000)
```

```{r}
hist(mpsz$`PreSch Density`)
```

```{r}
ggplot(data=mpsz, 
       aes(x= as.numeric(`PreSch Density`)))+
  geom_histogram(bins=20, 
                 color="black", 
                 fill="light blue") +
  labs(title = "Are pre-school even distributed in Singapore?",
       subtitle= "There are many planning subzones with a single pre-school, on the other hand, \nthere are seven planning subzones with at least 30 or more pre-schools",
      x = "Pre-school density (per km sq)",
      y = "Frequency")
```

```{r}
ggplot(data=mpsz, 
       aes(y = `PreSch Count`, 
           x= as.numeric(`PreSch Density`)))+
  geom_point(color="black", 
             fill="light blue") +
  xlim(0, 40) +
  ylim(0, 40) +
  labs(title = "",
      x = "Pre-school density (per km sq)",
      y = "Pre-school count")
```
