---
title: "Hands-on Exercise 01"
author: "Kalpit"
---
## Introduction to Geospatial Data Science
Geospatial Data Science is a process of importing, wrangling, integrating, and processing geographically referenced data sets. In this hands-on exercise, you will learn how to perform geospatial data science tasks in R by using sf package.

```{r}
pacman::p_load(sf, tidyverse)
```

## Importing the data

### Importing Polygon feature data for Masterplan

```{r}
mpsz = st_read(dsn = "K:/kalpitkulshrestha24/ISSS626/Hands-on_Exercise/Hands-on_EX01/data/geospatial/MasterPlan2014SubzoneBoundaryWebSHP", 
                  layer = "MP14_SUBZONE_WEB_PL")

```

### Importing polyline feature data in shapefile form for Cycling

```{r}
Cyclingpath = st_read(dsn = "K:/kalpitkulshrestha24/ISSS626/Hands-on_Exercise/Hands-on_EX01/data/geospatial/CyclingPath_Apr2025", layer = "CyclingPathGazette")

```
### Importing polyline feature data in shapefile form for Preschoolslocation dataset

```{r}
preschool = st_read("K:/kalpitkulshrestha24/ISSS626/Hands-on_Exercise/Hands-on_EX01/data/geospatial/PreSchoolslocation/PreSchoolslocation.kml")
```
## A simple feature dataframe
### Retriving geometery list-column

```{r}
st_geometry(mpsz)
```

```{r}
st_geometry(preschool)
```

```{r}
st_geometry(Cyclingpath)
```
### Using glimpse() funtion

```{r}
glimpse(mpsz)
```

### Using head() funtion

```{r}
head(mpsz)
```
## ploting the Geospatial data
```{r}
plot(mpsz)
```









