---
title: "Hands-on EX06"
author: "Kalpit Kulshrestha"
date: "02/october/2024"
date-modified: "last-modified"
---

# Geographical Segmentation with Spatially Constrained Clustering Techniques

## Overview

in this exercise we learn how to delineate homogeneous region by using greographically referernced multivriate data. There are two major analysis are:

-   Hierarchical cluster analysis
-   spatially constrained cluster analysis

## Getting Started

### The analytical question

In geobusiness and spatial policy, it is a common practice to delineate the market or planning area into homogeneous regions by using multivariate data.we are interested to delineate Shan State, Myanmar into homogeneous regions by using multiple Information and Communication technology (ICT) measures, namely: Radio, Television, Land line phone, Mobile phone, Computer, and Internet at home.

## The data

two dataset will be used for this hands on exercise:

-   Mayanmar township boundary data : ESRI shapefile format, consist of township boundary information of myanmar.
-   shan-ICT.csv this is an extract of the 2014 Myanmar Population and Housing Census Myanmar at the township level.

### Installing and loading R packages

The R packages needed for this exercise are as follows:

-   Spatial data handling
    -   sf, rgdal and spdep
-   Attribute data handling
    -   tidyverse, especially readr, ggplot2 and dplyr
-   Choropleth mapping
    -   tmap
-   Multivariate data visualisation and analysis
    -   coorplot, ggpubr, and heatmaply
-   Cluster analysis
    -   cluster
    -   ClustGeo
    
```{r}
pacman::p_load(spdep, tmap, sf, ClustGeo, 
               ggpubr, cluster, factoextra, NbClust,
               heatmaply, corrplot, psych, tidyverse, GGally)
```    
## Data import and preparation

### Importing Geospatial data into R environment
The Myanmar Township Boundary GIS data is in ESRI shapefile format. It will be imported into R environment by using the st_read() function of sf.
```{r}
shan_sf <- st_read(dsn = "data/geospatial", 
                   layer = "myanmar_township_boundaries") %>%
  filter(ST %in% c("Shan (East)", "Shan (North)", "Shan (South)")) %>%
  select(c(2:7))
```
The imported township boundary object is called shan_sf. It is saved in simple feature data.frame format. We can view the content of the newly created shan_sf simple features data.frame by using the code chunk below.
```{r}
shan_sf
```
Since shan_sf is conformed to tidy framework, we can also glimpse() to reveal the data type of it’s fields.
```{r}
glimpse(shan_sf)
```
### Importing aspatial data into R environment
```{r}
ict <- read_csv("data/aspatial/shan-ICT.csv")
```
```{r}
summary(ict)
```
### Derive new variables using dplur packages
The unit of measurement of the values are number of household. Using these values directly will be bias by the underlying total number of households. In general, the townships with relatively higher total number of households will also have higher number of households owning radio, TV, etc.

In order to overcome this problem, we will derive the penetration rate of each ICT variable by using the code chunk below.

```{r}
ict_derived <- ict %>%
  mutate(`RADIO_PR` = `Radio`/`Total households`*1000) %>%
  mutate(`TV_PR` = `Television`/`Total households`*1000) %>%
  mutate(`LLPHONE_PR` = `Land line phone`/`Total households`*1000) %>%
  mutate(`MPHONE_PR` = `Mobile phone`/`Total households`*1000) %>%
  mutate(`COMPUTER_PR` = `Computer`/`Total households`*1000) %>%
  mutate(`INTERNET_PR` = `Internet at home`/`Total households`*1000) %>%
  rename(`DT_PCODE` =`District Pcode`,`DT`=`District Name`,
         `TS_PCODE`=`Township Pcode`, `TS`=`Township Name`,
         `TT_HOUSEHOLDS`=`Total households`,
         `RADIO`=`Radio`, `TV`=`Television`, 
         `LLPHONE`=`Land line phone`, `MPHONE`=`Mobile phone`,
         `COMPUTER`=`Computer`, `INTERNET`=`Internet at home`) 

summary(ict_derived)
```
:::callout-note
Notice that six new fields have been added into the data.frame. They are RADIO_PR, TV_PR, LLPHONE_PR, MPHONE_PR, COMPUTER_PR, and INTERNET_PR.
:::

## Exploratory data analysis (EDA)

### EDA using statitical graphics
Histogram is useful to identify the overall distribution of the data values (i.e. left skew, right skew or normal distribution)

```{r}
ggplot(data=ict_derived, 
       aes(x=`RADIO`)) +
  geom_histogram(bins=20, 
                 color="black", 
                 fill="light blue")
```
** Boxplot is used to detect outliers in the data**
```{r}
ggplot(data=ict_derived, 
       aes(x=`RADIO`)) +
  geom_boxplot(color="black", 
               fill="light blue")
```
we will also plotting the distribution of the newly derived variables (i.e. Radio penetration rate) by using the code chunk below.
```{r}
ggplot(data=ict_derived, 
       aes(x=`RADIO_PR`)) +
  geom_histogram(bins=20, 
                 color="black", 
                 fill="light blue")
```
```{r}
ggplot(data=ict_derived, 
       aes(x=`RADIO_PR`)) +
  geom_boxplot(color="black", 
               fill="light blue")
```
```{r}
radio <- ggplot(data=ict_derived, 
             aes(x= `RADIO_PR`)) +
  geom_histogram(bins=20, 
                 color="black", 
                 fill="light blue")

tv <- ggplot(data=ict_derived, 
             aes(x= `TV_PR`)) +
  geom_histogram(bins=20, 
                 color="black", 
                 fill="light blue")

llphone <- ggplot(data=ict_derived, 
             aes(x= `LLPHONE_PR`)) +
  geom_histogram(bins=20, 
                 color="black", 
                 fill="light blue")

mphone <- ggplot(data=ict_derived, 
             aes(x= `MPHONE_PR`)) +
  geom_histogram(bins=20, 
                 color="black", 
                 fill="light blue")

computer <- ggplot(data=ict_derived, 
             aes(x= `COMPUTER_PR`)) +
  geom_histogram(bins=20, 
                 color="black", 
                 fill="light blue")

internet <- ggplot(data=ict_derived, 
             aes(x= `INTERNET_PR`)) +
  geom_histogram(bins=20, 
                 color="black", 
                 fill="light blue")
```

`ggarrange()` function of `ggpubr` packages is used to group these histogramns together.
```{r}
ggarrange(radio, tv, llphone, mphone, computer, internet, 
          ncol = 3, 
          nrow = 2)
```
### EDA using Choropleth map

#### Joining Geospatial and asptial data
Before we can prepare the choropleth map, we need to combine both the geospatial data object (i.e. shan_sf) and aspatial data.frame object (i.e. ict_derived) into one. This will be performed by using the left_join function of dplyr package. The shan_sf simple feature data.frame will be used as the base data object and the ict_derived data.frame will be used as the join table.
```{r}
shan_sf <- left_join(shan_sf, 
                     ict_derived, by=c("TS_PCODE"="TS_PCODE"))
  
write_rds(shan_sf, "data/rds/shan_sf.rds")
```

```{r}
shan_sf <- read_rds("data/rds/shan_sf.rds")
```

#### Preparing a choropleth map
To have a quick look at the distribution of Radio penetration rate of Shan State at township level, a choropleth map will be prepared.

we use `qtm()` function of **tmap** package.
```{r}
qtm(shan_sf,"RADIO_PR")
```
In order to reveal the distribution shown in the choropleth map above are bias to the underlying total number of households at the townships, we will create two choropleth maps, one for the total number of households (i.e. TT_HOUSEHOLDS.map) and one for the total number of household with Radio (RADIO.map) by using the code chunk below.
```{r}
TT_HOUSEHOLDS.map <- tm_shape(shan_sf) + 
  tm_fill(col = "TT_HOUSEHOLDS",
          n = 5,
          style = "jenks", 
          title = "Total households") + 
  tm_borders(alpha = 0.5) 

RADIO.map <- tm_shape(shan_sf) + 
  tm_fill(col = "RADIO",
          n = 5,
          style = "jenks",
          title = "Number Radio ") + 
  tm_borders(alpha = 0.5) 

tmap_arrange(TT_HOUSEHOLDS.map, RADIO.map,
             asp=NA, ncol=2)
```
Notice that the choropleth maps above clearly show that townships with relatively larger number ot households are also showing relatively higher number of radio ownership.

Now let us plot the choropleth maps showing the dsitribution of total number of households and Radio penetration rate by using the code chunk below.
```{r}
tm_shape(shan_sf) +
    tm_polygons(c("TT_HOUSEHOLDS", "RADIO_PR"),
                style="jenks") +
    tm_facets(sync = TRUE, ncol = 2) +
  tm_legend(legend.position = c("right", "bottom"))+
  tm_layout(outer.margins=0, asp=0)
```




