---
title: "Hands-on EX08"
author: "Kalpit Kulshrestha"
---
# Geographically Weighted Predictive Models

## overview
Predictive modelling uses statistical learning or machine learning techniques to predict outcomes. By and large, the event one wants to predict is in the future. However, a set of known outcome and predictors (also known as variables) will be used to calibrate the predictive models.

Geospatial predictive modelling is conceptually rooted in the principle that the occurrences of events being modeled are limited in distribution. When geographically referenced data are used, occurrences of events are neither uniform nor random in distribution over space. There are geospatial factors (infrastructure, sociocultural, topographic, etc.) that constrain and influence where the locations of events occur. Geospatial predictive modeling attempts to describe those constraints and influences by spatially correlating occurrences of historical geospatial locations with environmental factors that represent those constraints and influences.

### Learning outcome
In this in-class exercise, you will learn how to build predictive model by using geographical random forest method. By the end of this hands-on exercise, you will acquire the skills of:

preparing training and test data sets by using appropriate data sampling methods,
calibrating predictive models by using both geospatial statistical learning and machine learning methods,
comparing and selecting the best model for predicting the future outcome,
predicting the future outcomes by using the best model calibrated.

## Installing and Loading R packages
```{r}
pacman::p_load(sf, spdep, GWmodel, SpatialML, 
               tmap, rsample, Metrics, tidyverse)
```

## Preparing Data

```{r}
mdata <- read_rds("data/rawdata/mdata.rds")
```
### Data Sampling
The entire data are split into training and test data sets with 65% and 35% respectively by using initial_split() of rsample package. rsample is one of the package of tigymodels.

```{r}
set.seed(1234)
resale_split <- initial_split(mdata, 
                              prop = 6.5/10,)
train_data <- training(resale_split)
test_data <- testing(resale_split)
```

```{r}
write_rds(train_data, "data/model/train_data.rds")
write_rds(test_data, "data/model/test_data.rds")
```
## Computing Correlation Matrix
```{r}
mdata_nogeo <- mdata %>%
  st_drop_geometry()
corrplot::corrplot(cor(mdata_nogeo[, 2:17]), 
                   diag = FALSE, 
                   order = "AOE",
                   tl.pos = "td", 
                   tl.cex = 0.5, 
                   method = "number", 
                   type = "upper")
```
## Retriving the Stored Data
```{r}
train_data <- read_rds("data/model/train_data.rds")
test_data <- read_rds("data/model/test_data.rds")
```

## Building a non-spatial multiple linear regression
```{r}
price_mlr <- lm(resale_price ~ floor_area_sqm +
                  storey_order + remaining_lease_mths +
                  PROX_CBD + PROX_ELDERLYCARE + PROX_HAWKER +
                  PROX_MRT + PROX_PARK + PROX_MALL + 
                  PROX_SUPERMARKET + WITHIN_350M_KINDERGARTEN +
                  WITHIN_350M_CHILDCARE + WITHIN_350M_BUS +
                  WITHIN_1KM_PRISCH,
                data=train_data)
summary(price_mlr)
```
```{r}
write_rds(price_mlr, "data/model/price_mlr.rds" ) 
```

## gwr predictive model

### Computing adaptive bandwidth
```{r}
bw_adaptive <- bw.gwr(
  formula = resale_price ~ floor_area_sqm +
    storey_order + remaining_lease_mths +
    PROX_CBD + PROX_ELDERLYCARE + PROX_HAWKER +
    PROX_MRT + PROX_PARK + PROX_MALL + 
    PROX_SUPERMARKET + WITHIN_350M_KINDERGARTEN +
    WITHIN_350M_CHILDCARE + WITHIN_350M_BUS +
    WITHIN_1KM_PRISCH,
  data=train_data,
  approach="CV",
  kernel="gaussian",
  adaptive=TRUE,
  longlat=FALSE)

write_rds(bw_adaptive, "data/model/bw_adaptive.rds")
```

### Constructing the adaptive bandwidth gwr model
First, let us call the save bandwidth by using the code chunk below.
```{r}
bw_adaptive <- read_rds(
  "data/model/bw_adaptive.rds")
```
Now, we can go ahead to calibrate the gwr-based hedonic pricing model by using adaptive bandwidth and Gaussian kernel as shown in the code chunk below.
```{r}
gwr_adaptive <- gwr.basic(
  formula = resale_price ~ floor_area_sqm +
    storey_order + remaining_lease_mths + 
    PROX_CBD + PROX_ELDERLYCARE + 
    PROX_HAWKER + PROX_MRT + PROX_PARK + 
    PROX_MALL + PROX_SUPERMARKET +
    WITHIN_350M_KINDERGARTEN +
    WITHIN_350M_CHILDCARE + 
    WITHIN_350M_BUS + WITHIN_1KM_PRISCH,
  data=train_data,
  bw=bw_adaptive, 
  kernel = 'gaussian', 
  adaptive=TRUE,
  longlat = FALSE)

write_rds(gwr_adaptive, 
          "data/model/gwr_adaptive.rds")
```

### Retrieve gwr output object
The code chunk below will be used to retrieve the save gwr model object.
```{r}
gwr_adaptive <- read_rds("data/model/gwr_adaptive.rds")
```
The code below can be used to display the model output.
```{r}
gwr_adaptive
```
### Computing adaptive bandwidth for the test data
```{r}
gwr_bw_test_adaptive <- bw.gwr(
  formula = resale_price ~ floor_area_sqm +
    storey_order + remaining_lease_mths +
    PROX_CBD + PROX_ELDERLYCARE + PROX_HAWKER +
    PROX_MRT + PROX_PARK + PROX_MALL + 
    PROX_SUPERMARKET + WITHIN_350M_KINDERGARTEN +
    WITHIN_350M_CHILDCARE + WITHIN_350M_BUS +
    WITHIN_1KM_PRISCH,
  data=test_data,
  approach="CV",
  kernel="gaussian",
  adaptive=TRUE,
  longlat=FALSE)

write_rds(gwr_bw_test_adaptive, 
          "data/model/gwr_bw_test_adaptive.rds")
```

### Computing predicted values of the test data
```{r}
gwr_pred <- gwr.predict(
  formula = resale_price ~ floor_area_sqm + 
    storey_order + remaining_lease_mths + 
    PROX_CBD + PROX_ELDERLYCARE + PROX_HAWKER + 
    PROX_MRT + PROX_PARK + PROX_MALL + 
    PROX_SUPERMARKET + WITHIN_350M_KINDERGARTEN +
    WITHIN_350M_CHILDCARE + WITHIN_350M_BUS + 
    WITHIN_1KM_PRISCH, 
  data=train_data, 
  predictdata = test_data, 
  bw=40, 
  kernel = 'gaussian', 
  adaptive=TRUE, 
  longlat = FALSE)

write_rds(gwr_pred, 
          "data/model/gwr_pred.rds")
```

## Calibrating Random Forest Model
In this section, you will learn how to calibrate a model to predict HDB resale price by using random forest function of ranger package.

### Preparing coordinates data
The code chunk below extract the x,y coordinates of the full, training and test data sets.
```{r}
coords <- st_coordinates(mdata)
coords_train <- st_coordinates(train_data)
coords_test <- st_coordinates(test_data)
```

Before continue, we write all the output into rds for future used.
```{r}
coords_train <- write_rds(
  coords_train, 
  "data/model/coords_train.rds" )
coords_test <- write_rds(
  coords_test, 
  "data/model/coords_test.rds" )
```
### Droping geometry field
First, we will drop geometry column of the sf data.frame by using st_drop_geometry() of sf package.
```{r}
train_data_nogeom <- train_data %>% 
  st_drop_geometry()

test_data_nogeom <- test_data %>% 
  st_drop_geometry()
```
### Calibrating a non-spatial random forest model
```{r}
set.seed(1234)
rf <- ranger(
  formula = resale_price ~ floor_area_sqm +
    storey_order + remaining_lease_mths + 
    PROX_CBD + PROX_ELDERLYCARE + 
    PROX_HAWKER + PROX_MRT + PROX_PARK + 
    PROX_MALL + PROX_SUPERMARKET +
    WITHIN_350M_KINDERGARTEN + 
    WITHIN_350M_CHILDCARE + 
    WITHIN_350M_BUS +  
    WITHIN_1KM_PRISCH,
  data=train_data_nogeom)
rf
```

## Calibrating Geographical Weighted Random Forest Model
In this section, you will learn how to calibrate a model to predict HDB resale price by using grf() of SpatialML package.

### Calibrating using training data
The code chunk below calibrate a geographic ranform forest model by using grf() of SpatialML package.
```{r}
set.seed(1234)
gwRF_adaptive <- grf(
  formula = resale_price ~ floor_area_sqm +
    storey_order + remaining_lease_mths + 
    PROX_CBD + PROX_ELDERLYCARE + 
    PROX_HAWKER + PROX_MRT + PROX_PARK + 
    PROX_MALL + PROX_SUPERMARKET +
    WITHIN_350M_KINDERGARTEN +
    WITHIN_350M_CHILDCARE + 
    WITHIN_350M_BUS + WITHIN_1KM_PRISCH,
  dframe=train_data_nogeom, 
  bw=40,
  kernel="adaptive",
  coords=coords_train)

write_rds(gwRF_adaptive, "data/model/gwRF_adaptive.rds")
```

```{r}
gwRF_adaptive <- read_rds("data/model/gwRF_adaptive.rds")
```


### Predicting by using test data
#### Preparing the test data
The code chunk below will be used to combine the test data with its corresponding coordinates data.
```{r}
test_data_nogeom <- cbind(
  test_data, coords_test) %>%
  st_drop_geometry()
```
#### Predicting with test data
Next, predict.grf() of spatialML package will be used to predict the resale value by using the test data and gwRF_adaptive model calibrated earlier.
```{r}
gwRF_pred <- predict.grf(gwRF_adaptive, 
                           test_data_nogeom, 
                           x.var.name="X",
                           y.var.name="Y", 
                           local.w=1,
                           global.w=0)

GRF_pred <- write_rds(gwRF_pred, "data/model/GRF_pred.rds")
```
#### Converting the predicting output into a data frame
The output of the predict.grf() is a vector of predicted values. It is wiser to convert it into a data frame for further visualisation and analysis.
```{r}
GRF_pred <- read_rds(
  "chap14/data/model/GRF_pred.rds")
GRF_pred_df <- as.data.frame(GRF_pred)
```
In the code chunk below, cbind() is used to append the predicted values onto test_datathe
```{r}
test_data_p <- cbind(
  test_data, GRF_pred_df) %>%
  select(resale_price, GRF_pred)

write_rds(test_data_p, 
          "chap14/data/model/test_data_p.rds")
```

### Calculating Root Mean Square Error
The root mean square error (RMSE) allows us to measure how far predicted values are from observed values in a regression analysis. In the code chunk below, rmse() of Metrics package is used to compute the RMSE.
```{r}
rmse(test_data_p$resale_price, 
     test_data_p$GRF_pred)
```

### Visualising the predicted values
Alternatively, scatterplot can be used to visualise the actual resale price and the predicted resale price by using the code chunk below.
```{r}
ggplot(data = test_data_p,
       aes(x = GRF_pred,
           y = resale_price)) +
  geom_point()
```
























